---
title: "DS 311 group project"
author: "Pyi Than"
date: "2022-11-09"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Importing library

```{r}
library(readxl)
library(dplyr)
library(tidyverse)
library(ggplot2)
```
## Importing data

```{r}
salary=data.frame(read_excel("salary_data_states.xlsx"))
```
## Look at column name

```{r}
colnames(salary)
```

## Check the data type

```{r}
str(salary)
```

## Check Unique job title sub group

```{r}
unique(salary$JOB_TITLE_SUBGROUP)
```
## Check Unique work state

```{r}
unique(salary$WORK_STATE)
```

## Check missing values

```{r}
apply(X=is.na(salary),MARGIN=2,FUN=sum)
```

## What states(of those I am willing to move to) have the highest paying data-related salaries?

Average data scientist salary by America. 
```{r}
salary_dataScientist=salary%>%
  group_by(WORK_STATE,JOB_TITLE_SUBGROUP)%>%
  filter(JOB_TITLE_SUBGROUP %in% c("data scientist"))%>%
  summarize(AverageAnnualSalary=mean(PAID_WAGE_PER_YEAR))%>%
  arrange(-AverageAnnualSalary)
salary_dataScientist
write.csv(salary_dataScientist,'C:\\Users\\Aungkyaw\\Desktop\\salary_dataScientist.csv')
```
The state I choose is California, Washington, Hawaii, Wisconsin and New Mexico
## Average data scientist salary by state.
```{r}
## data scientist
salary_dataScientist_state=salary%>%
  group_by(WORK_STATE,JOB_TITLE_SUBGROUP)%>%
  filter(JOB_TITLE_SUBGROUP %in% c("data scientist"))%>%
  filter(WORK_STATE %in% c("California","Washington","Hawaii","Wisconsin","New Mexico"))%>%
  summarize(AverageAnnualSalary=mean(PAID_WAGE_PER_YEAR))%>%
  arrange(-AverageAnnualSalary)
salary_dataScientist_state

# California has the highest average annual salary for data scientist.Follow by Washington and Minnesota. On the other hand, Florida has the lowest average annual salary for data scientist.
```

## Which companies have the highest salaries for data scientist in California

```{r}
salary_dataScientist_company_CA=salary%>%
  group_by(WORK_STATE,EMPLOYER_NAME,JOB_TITLE_SUBGROUP)%>%
   filter(JOB_TITLE_SUBGROUP %in% c("data scientist"))%>%
  filter(WORK_STATE %in% c("California"))%>%
  summarize(AverageAnnualSalary=mean(PAID_WAGE_PER_YEAR))%>%
  arrange(-AverageAnnualSalary)%>%
  head(5)
salary_dataScientist_company_CA
write.csv(salary_dataScientist_company_CA,'C:\\Users\\Aungkyaw\\Desktop\\salary_dataScientist_company_CA.csv')
```
## Which companies have the highest salaries for data scientist in Washington

```{r}
salary_dataScientist_company_WA=salary%>%
  group_by(WORK_STATE,EMPLOYER_NAME,JOB_TITLE_SUBGROUP)%>%
   filter(JOB_TITLE_SUBGROUP %in% c("data scientist"))%>%
  filter(WORK_STATE %in% c("Washington"))%>%
  summarize(AverageAnnualSalary=mean(PAID_WAGE_PER_YEAR))%>%
  arrange(-AverageAnnualSalary)%>%
  head(5)
salary_dataScientist_company_WA
write.csv(salary_dataScientist_company_WA,'C:\\Users\\Aungkyaw\\Desktop\\salary_dataScientist_company_WA.csv')
```

## Average data Analyst salary in America. 

```{r}
salary_dataAnalyst=salary%>%
  group_by(WORK_STATE,JOB_TITLE_SUBGROUP)%>%
  filter(JOB_TITLE_SUBGROUP %in% c("data analyst"))%>%
  summarize(AverageAnnualSalary=mean(PAID_WAGE_PER_YEAR))%>%
  arrange(-AverageAnnualSalary)
salary_dataAnalyst
write.csv(salary_dataAnalyst,'C:\\Users\\Aungkyaw\\Desktop\\salary_dataAnalysst.csv')
```

## Average data analyst salary by state

```{r}
salary_dataAnalyst_state=salary%>%
  group_by(WORK_STATE,JOB_TITLE_SUBGROUP)%>%
  filter(JOB_TITLE_SUBGROUP %in% c("data analyst"))%>%
  filter(WORK_STATE %in% c("California","Washington","Hawaii","Wisconsin","New Mexico"))%>%
  summarize(AverageAnnualSalary=mean(PAID_WAGE_PER_YEAR))%>%
  arrange(-AverageAnnualSalary)
salary_dataAnalyst_state
```

## Average business Analyst salary in America. 

```{r}
salary_businessAnalyst=salary%>%
  group_by(WORK_STATE,JOB_TITLE_SUBGROUP)%>%
  filter(JOB_TITLE_SUBGROUP %in% c("business analyst"))%>%
  summarize(AverageAnnualSalary=mean(PAID_WAGE_PER_YEAR))%>%
  arrange(-AverageAnnualSalary)
salary_businessAnalyst
write.csv(salary_businessAnalyst,'C:\\Users\\Aungkyaw\\Desktop\\salary_businessAnalysst.csv')
```

## Average business analyst salary by state

```{r}
salary_businessAnalyst_state=salary%>%
  group_by(WORK_STATE,JOB_TITLE_SUBGROUP)%>%
  filter(JOB_TITLE_SUBGROUP %in% c("business analyst"))%>%
  filter(WORK_STATE %in% c("California","Washington","Hawaii","Wisconsin","New Mexico"))%>%
  summarize(AverageAnnualSalary=mean(PAID_WAGE_PER_YEAR))%>%
  arrange(-AverageAnnualSalary)
salary_businessAnalyst_state
write.csv(salary_businessAnalyst_state,'C:\\Users\\Aungkyaw\\Desktop\\salary_businessAnalyst_state.csv')
```

## Which company gives highest average data scientist salary in CA

```{r}
salary_dataScientist_company_ca=salary%>%
  group_by(WORK_STATE,EMPLOYER_NAME,JOB_TITLE_SUBGROUP)%>%
  filter(JOB_TITLE_SUBGROUP %in% c("data scientist"))%>%
  filter(WORK_STATE %in% c("California"))%>%
  summarize(AverageAnnualSalary=mean(PAID_WAGE_PER_YEAR))%>%
  arrange(-AverageAnnualSalary)
salary_dataScientist_company_ca
write.csv(salary_dataScientist_company_ca,'C:\\Users\\Aungkyaw\\Desktop\\salary_dataScientist_company_ca.csv')
```

## Which company gives highest average data scientist salary in WA

```{r}
salary_dataScientist_company_wa=salary%>%
  group_by(WORK_STATE,EMPLOYER_NAME,JOB_TITLE_SUBGROUP)%>%
  filter(JOB_TITLE_SUBGROUP %in% c("data scientist"))%>%
  filter(WORK_STATE %in% c("Washington"))%>%
  summarize(AverageAnnualSalary=mean(PAID_WAGE_PER_YEAR))%>%
  arrange(-AverageAnnualSalary)
salary_dataScientist_company_wa
write.csv(salary_dataScientist_company_wa,'C:\\Users\\Aungkyaw\\Desktop\\salary_dataScientist_company_wa.csv')
```
## Which company gives highest average business analyst salary in CA

```{r}
salary_businessAnalyst_company_ca=salary%>%
  group_by(WORK_STATE,EMPLOYER_NAME,JOB_TITLE_SUBGROUP)%>%
  filter(JOB_TITLE_SUBGROUP %in% c("business analyst"))%>%
  filter(WORK_STATE %in% c("California"))%>%
  summarize(AverageAnnualSalary=mean(PAID_WAGE_PER_YEAR))%>%
  arrange(-AverageAnnualSalary)%>%
  head(5)
salary_businessAnalyst_company_ca
write.csv(salary_businessAnalyst_company_ca,'C:\\Users\\Aungkyaw\\Desktop\\salary_businessAnalyst_company_ca.csv')
```

## Which company gives highest average business analyst salary in WA

```{r}
salary_businessAnalyst_company_wa=salary%>%
  group_by(WORK_STATE,EMPLOYER_NAME,JOB_TITLE_SUBGROUP)%>%
  filter(JOB_TITLE_SUBGROUP %in% c("business analyst"))%>%
  filter(WORK_STATE %in% c("Washington"))%>%
  summarize(AverageAnnualSalary=mean(PAID_WAGE_PER_YEAR))%>%
  arrange(-AverageAnnualSalary)%>%
  head(5)
salary_businessAnalyst_company_wa
write.csv(salary_businessAnalyst_company_wa,'C:\\Users\\Aungkyaw\\Desktop\\salary_businessAnalyst_company_wa.csv')
```
